Global optimization method with dual Lipschitz constant estimates for problems with non-convex constraints

### 2020-01-28 2019/NUMTA_Barkalov/Special Issure/article.tex
New results for Simple_4 class
Last corrections before submitting

	It is supposed that the objective function $\varphi(y)$ (henceforth denoted by $g_{m+1}(y)$) and the left-hand sides $g_{j}(y), 1 \leq j \leq m$, of the constraints satisfy the Lipschitz condition
\begin{equation}\label{lipschitz_condition}
	\left|g_j(y')-g_j (y'')\right| \leq L_j \|y'-y''\|, \; y', y''\in D, \; 1\leq j \leq m+1.
\end{equation}
with a priori unknown constants $L_j, 1\leq j \leq m+1$, and may be multiextremal.

@@

	The existence of the $\varepsilon$-reserved solutions of problem (\ref{problem}) can be interpreted as some analog of the regularity condition in classical nonlinear programming problems. The applied role of this condition should be also noted. Even if the exact solution $y^\ast$ is found, its practical implementation is possible as some approximation of $y^\ast$ only. Therefore, it is important that there are feasible points from $Q_{m+1}$ close enough to $y^\ast$ (in coordinates and values). The existence of the $\varepsilon$-reserved solution guarantees the existence of such points. In this case, the domain $Y_{\epsilon}$ may play the role of a set of suitable approximations.

@@

	Thus, a search trial at some point $x^k \in [0,1]$ performed at the $k^{th}$ iteration of the algorithm will consist of the following sequence of operations:
\begin{enumerate} 
  \item Determine the image $y^k=y(x^k )$ with the use of the evolvent $y(x)$. 
  \item Compute the values $g_1 (y^k ), \ldots ,g_{\nu} (y^k )$ where $\nu = \nu(x)$ is the index of the point.
	\item Fix the result of the trial as three values 
\end{enumerate}

\begin{equation}\label{trial_result}
	x^k, \; z^k = g_{\nu}\left( y(x^k) \right), \; \nu = \nu(x^k),
\end{equation}

@@

	In the case when the point $y^k = y(x^k)$ is a feasible one, i.e. when $x^k \in Q_{m+1}$, the trial involves checking all constraints and computing the objective function. Here, the value of the index is accepted to be equal to the quantity $\nu = m+1$.

	The main idea of the index scheme of accounting for the constraints consists in the reduction of the constrained problem (\ref{one_dimensional_problem}) (the functions of which satisfy the H\"{o}lder condition with corresponding coefficients $K_j$) to an unconstrained problem of the form
\begin{equation}\label{reduction_problem}
	\psi(x^*)=\min \left\{\psi(x): x \in [0,1] \right\},
\end{equation}
\begin{eqnarray} \nonumber
	\psi(x)=\frac{g_{\nu}(y(x))}{K_{\nu}} - 
	\left\{
   \begin{array}{lr}
     0, & \nu \leq m,\\
     \frac{\varphi^\ast}{K_{m+1}}, & \nu = m + 1.
   \end{array}
	\right.
\end{eqnarray}
where $\nu = \nu(x)$ and $\varphi^\ast$ is the objective function value at the point of the problem solution (\ref{one_dimensional_problem}).

@@

\section{Index-Based Global Search Algorithm}
	The index-based global search algorithm under consideration for solving the problem (\ref{problem}) (here it is given according to \cite{Strongin2000}) involves the construction of a sequence of points $x^k \in (0,1)$, at which the search trials are conducted with the results of the type (\ref{trial_result}). The set of triplets $\{x^k, z^k=g_{\nu}\big(y(x^k)\big), \nu=\nu(x^k)\}, 1 \leq k \leq n,$  makes up the search information accumulated by the method after performing $n$ steps. The operation of the index-based global search algorithm is defined by the following rules.

@@

	The convergence conditions for the index global search algorithm can be formulated in the form of the following theorem from \cite{Strongin2000}. 
\\
\textbf{Theorem}. Assume that the following conditions are satisfied:
\begin{enumerate} 
  \item Problem (\ref{problem}) has an $\varepsilon$-reserved solution $y_{\varepsilon}$ from (\ref{epsilon_reserved_solution}).
  \item Functions $g_j(y), 1 \leq j \leq m+1,$ admit Lipschitzian (with the constants $L_j$) extensions $G_{j}(y)$ over the whole feasible domain $D$ from (\ref{D}), i.e.
$$
	g_j \left( y(x) \right) = G_j \left( y(x) \right), \; x \in Q_j, \; 1 \leq j \leq m+1,
$$
where the sets $Q_j$ are from (\ref{Q_intervals}), and $y(x)$ is the curve from (\ref{one_dimensional_problem}).
	\item The parameters of the method $\epsilon_{\nu}, 1 \leq \nu \leq m,$ used in the rule (\ref{z_estimates}) are the corresponding components of the reserve vector $\epsilon_{R}$ from (\ref{epsilon_reserved_solution}).
	\item For a sufficiently large iteration number $k$, the values $\mu_{\nu}$ from (\ref{current_lower_bounds}) satisfy the inequalities
		\begin{equation}\label{theorem_inequalities}
			r_{\nu}\mu_{\nu} > 2^{3-\frac{1}{N}}L_{\nu}\sqrt{N+3}, \; 1 \leq \nu \leq m+1. 
		\end{equation}
Then, any limit point $\bar{y}$ of the sequence of the trials $\{y^k\}=\big\{y\big(x^k\big)\big\}$ generated by the index algorithm for problem (\ref{problem}) belongs to the set $Y_{\epsilon}$ form (\ref{Y_epsilon}) and satisfies the conditions	
$$
	\varphi(\bar y) = \inf\left\{\varphi(y^k):g_j(y^k) \leq 0, 1 \leq j \leq m, k = 1,2,\ldots \right\} \leq \varphi(y_{\epsilon}).
$$
\end{enumerate}
\textbf{Note 1}. The computational scheme of the algorithm is described in the form convenient for  formulating its theoretical properties. The software implementation of this scheme may be made more frugal. For example, according to Step 7, the point $x^{k+1}$ of the next trial will be an internal point of the interval $(x_{t-1}, x_t)$, i.e. $x^{k+1} \in (x_{t-1}, x_t)$. Therefore, the set of trial points $x^1,\ldots, x^k$ (and the related trial results (\ref{trial_result})) may not be ordered  at every step of the algorithm but may be maintained in an ordered state using appropriate data structures.
\\
\textbf{Note 2}. The use of the set of parameters $\epsilon_1,\ldots,\epsilon_m$ (the values of which correspond to the coordinates of the reserve vector $\epsilon_R$) in the index algorithm will significantly reduce the density of the trial sequence in the vicinity of the feasible domain boundaries. However, this raises a question about the particular values of the reserve vector components $\epsilon_R$. In \cite{Strongin2000}, it was proposed to choose the components of the vector $\epsilon_R$ according to the formula 
\begin{equation}\label{epsilon_nu}
	\epsilon_{\nu} = \mu_{\nu}\delta, \; 1 \leq \nu \leq m, 
\end{equation}
where $\mu_{\nu}$ are the current estimates from (\ref{current_lower_bounds}) and $\delta>0$ is a parameter. This approach to choosing the reserve vector components allows one to take into account the properties of different problem functions characterized by their Lipschitz constants. This method for choosing the reserve vector was used in the numerical experiments, the results of which are presented below.
\\
\textbf{Note 3}. The index algorithm for solving the constrained problems is based on the information-statistical algorithm of global search (AGS) for solving unconstrained problems. AGS has been compared repeatedly with both deterministic and nature-inspired global optimization algorithms. The results of the experimental comparison (see, for example, \cite{Sovrasov2019}) show that AGS is superior to many known methods of similar purpose, including DIRECT and DIRECT\textit{l}.

	At the same time, the index scheme of accounting for the constraints also demonstrated its advantages (in the number of computed values of the problem functions) over the traditional penalty functions method in solving a wide range of model problems (see the results of the experimental comparison in \cite{Barkalov2017_1, Barkalov2017_2}.)

@@

	According to Theorem 1 from Section 3, the sequence of trial points $\{y^k\}$ will converge to a global minimizer $y^\ast$ if the condition (\ref{theorem_inequalities}) is satisfied. So far, an appropriate choice of the parameters $r_{\nu}$ allows using the values
\begin{equation}\label{estimates_Lipschitz_constants}
	\frac{r_{\nu}\mu_{\nu}}{2^{3-\frac{1}{N}}\sqrt{N+3}}, \; 1 \leq \nu \leq m+1, 
\end{equation}
as the estimates of the Lipschitz constants for the problem functions $g_{j}(y), 1 \leq j \leq m+1$.

	Compliance with condition (\ref{theorem_inequalities}) will be achieved when choosing a high value of the parameters $r_{\nu}$. However, in this case the constant will be overestimated, and the method will perform a large number of trials until the stopping condition is fulfilled.
	The choice of low values of the parameters $r_{\nu}$ (which corresponds to an underestimate of the Lipschitz constant) would reduce the number of trials considerably but could destroy the convergence to the global extremum (due to possible violation of the condition (\ref{theorem_inequalities})).

@@

	Step 5. For each interval $(x_{i-1}, x_i), 1 \leq i \leq k+1,$ 
\begin{flushleft}
	\begin{itemize}[\labelitemii]
		\item Compute the global characteristics $R_{glob}(i)$ according to formulae (\ref{R_1}), (\ref{R_2}), (\ref{R_3}) using the value $r_{\nu}=r_{\nu}^{glob}$.
		\item Compute the local characteristics $R_{loc}(i)$ according to formulae (\ref{R_1}), (\ref{R_2}), (\ref{R_3}) using the value $r_{\nu}=r_{\nu}^{loc}$.
		\item Determine the characteristic $R(i)$ as
\begin{equation}\label{R_max}
	R(i) = \max{\left\{ R_{glob}(i), \rho R_{loc}(i) \right\}},
\end{equation}
where $\rho=\Big(\frac{1-1/r_{M}^{glob}}{1-1/r_{M}^{loc}}\Big)^2$, if $\nu(x_{i-1})=\nu(x_i)=\nu$, and $\rho=1$, if $\nu(x_{i-1}) \neq \nu(x_i)$.
	\end{itemize}
\end{flushleft}

@@

	Let us illustrate the work of the algorithms IA and IA-DL when minimizing the function
\begin{eqnarray}
	\varphi(y_1, y_2)=-1.5y_1^2\exp{\left\{1-y_1^2-20.25(y_1-y_2)^2\right\}}- \nonumber \\
	-\left(0.5(y_1-1)(y_2-1)\right)^4\exp{\left\{2-\left(0.5(y_1-1)\right)^4-(y_2-1)^4\right\}}
	\nonumber
\end{eqnarray}
in the range $0 \leq y_1 \leq 4$, $1 \leq y_2 \leq 3$ subject to the constraints
\begin{eqnarray}
	g_1(y_1, y_2) &=& 0.01 \left( (y_1-2.2)^2+(y_2-1.2)^2-2.25 \right) \leq 0, \nonumber \\
	g_2(y_1, y_2) &=& 100 \left(1-(y_1-2)^2/1.44+(0.5y_2)^2 \right) \leq 0, \nonumber \\
	g_3(y_1, y_2) &=& 10 \left( y_2 - 1.5 - 1.5 \sin{\left( 6.283(y_1-1.75) \right)}\right) \leq 0. \nonumber
\end{eqnarray}
The feasible domain of this problem is not connected and consists of three subdomains. The global minimum $\varphi(y^\ast)=-1.489$ is achieved at the point $y^\ast =(0.942, 0.944)$ (the values are given with the accuracy of $10^{-3}$).

@@

	In Fig. \ref{fig:example_strongin}, the search domain (square) is shown, the feasible domain is highlighted by green, and the trial points generated by IA (a) and IA-DL (b) are marked. The best point is marked by red. For both methods, the evolvent with the parameter $m=10$ was used, the accuracy of the search for the solution $\epsilon=0.002$; the adaptive reserve parameter $\delta=0.008$ from (\ref{epsilon_nu}). For the IA method, the parameters $r_{\nu}=2.3, 1 \leq \nu \leq 4$ were used. The same values were used for the IA-DL method as the parameters $r_{\nu}^{glob}, 1 \leq \nu \leq 4$. Also, the parameters $r_{\nu}^{loc}=1.5, 1 \leq \nu \leq 4,$ were used. To solve the problem with a predefined accuracy $\epsilon=0.002$, 478 trials were required for the IA method compared to only 303 trials for the IA-DL method. The results of larger-scale numerical experiments are presented in Section 4.

	Let us formulate and prove the convergence conditions for the new algorithm IA-DL.\\
\textbf{Theorem 2}. Let us assume the index algorithm (IA) is applied to solve the problem (\ref{problem}) and for the values $r_{\nu}=r_{\nu}^{glob}, 1 \leq \nu \leq m+1,$ the conditions of Theorem 1 are fulfilled. Then, the algorithm with dual estimate of the Lipschitz constants (IA-DL) with the parameters $r_{\nu}^{glob}>r_{\nu}^{loc}>1, 1\leq \nu \leq m+1,$ when solving the same problem also generates a sequence of the trial points $\{y^k\}=\{y(x^k)\}$, any limit point $\bar{y}$ of which belongs to the set $Y_{\epsilon}$ from (\ref{Q_intervals}), i.e. satisfies the condition 
\begin{equation}\label{phi_bar_y}
	\varphi(\bar y) = \inf\left\{\varphi(y^k):g_j(y^k) \leq 0, 1 \leq j \leq m, k = 1,2,\ldots \right\} \leq \varphi(y_{\epsilon}).
\end{equation}

@@

	Since the limit point $\bar y = y(\bar x)$ belongs to the set $Y_{\epsilon}$, then
$$
	g_{\nu}\left( y(\bar x) \right) \leq -\epsilon_{\nu}, \; 1\leq\nu\leq m,
$$
$$
	g_{m+1}\left( y(\bar x) \right) \leq z_{m+1}^\ast.
$$

	Then, the characteristic $R_{glob}(j)$ of the interval $[x_{j-1}, x_j], j=j(k)$ will be strictly positive, i.e. $R_{glob}(j)>0$. Indeed, at $\nu=\nu(x_j)=\nu(x_{j-1})$ and the value $r_{\nu}=r_{\nu}^{glob}$
\begin{eqnarray}
	R(j) &=& \Delta_i + \frac{(z_j-z_{j-1})^2}{r_{\nu}^2\mu_{\nu}^2\Delta_i} - 2\frac{z_j+z_{j-1}-2z_{\nu}^*}{r_{\nu}\mu_{\nu}}  \nonumber \\
	&\geq& \Delta_i + 4\frac{z_{\nu}^*-\left( g_{\nu}\left( y(\bar x) \right)+L_{\nu}\sqrt{N+3}\left( (x_j-\bar x)^{\frac{1}{N}}+(\bar x - x_{j-1})^{\frac{1}{N}} \right)\right)}{r_{\nu}\mu_{\nu}} \nonumber \\
	&=& \Delta_i-4\frac{L_{\nu}\sqrt{N+3}\left( (x_j-\bar x)^{\frac{1}{N}}+(\bar x - x_{j-1})^{\frac{1}{N}} \right)}{r_{\nu}\mu_{\nu}}+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \nonumber \\
%	&=& \Delta_i-4\frac{L_{\nu}\sqrt{N+3}\left( \alpha^{\frac{1}{N}}+(1-\alpha)^{\frac{1}{N}} \right)}{r_{\nu}\mu_{\nu}}+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \nonumber \\
	&\geq& \Delta_i\left(1-4\frac{L_{\nu}\sqrt{N+3}\max_{0\leq\alpha\leq1} {\left( \alpha^{\frac{1}{N}}+(1-\alpha)^{\frac{1}{N}} \right)}}{r_{\nu}\mu_{\nu}} \right)+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \nonumber \\
	&=& \Delta_i\left(1-4\frac{2^{3-1/N}L_{\nu}\sqrt{N+3}}{r_{\nu}\mu_{\nu}} \right)+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \geq 0.
	\nonumber
\end{eqnarray}
At $\nu=\nu(x_j ) > \nu(x_{j-1})$ and the value $r_{\nu}=r_{\nu}^{glob}$
\begin{eqnarray}
	R(j) &=& 2\Delta_i - 4\frac{z_j-z_{\nu}^*}{r_{\nu}\mu_{\nu}} \geq 2\Delta_i+4\frac{z_{\nu}^*-\left( g_{\nu}\left(y(\bar x)\right)+2L_{\nu}\sqrt{N+3}(x_j-\bar x)^{\frac{1}{N}} \right)}{r_{\nu}\mu_{\nu}} \nonumber \\
	&=& 2\Delta_i - 8\frac{L_{\nu}\sqrt{N+3}(x_j-\bar x)^{\frac{1}{N}}}{r_{\nu}\mu_{\nu}}+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \nonumber \\
	&\geq& 2\Delta_i\left( 1-4\frac{L_{\nu}\sqrt{N+3}}{r_{\nu}\mu_{\nu}} \right) + 4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} > 0.
	\nonumber
\end{eqnarray}

@@

\begin{table}[h!]
	\centering
		\begin{tabular}{ccccc}
			\hline
			& $N=2$ & $N=3$ & $N=4$ & $N=5$ \\
			\hline
			IA-DL, $r_{\nu}^{loc}=1.4$ & 316 & 7403 & 44304 & 296357 \\
			\hline
			IA-DL, $r_{\nu}^{loc}=2.0$ & 350 & 8990(1) & 47133 & 360497 \\
			\hline
			IA & 439 & 11804(1) & 42837 & 404660(2) \\
			\hline
		\end{tabular}
	\caption{Average number of iterations performed by the methods on “Simple” class}
	\label{tab:average_number_of_iterations_simple}
\end{table}

>>>>

	It is supposed that the objective function $\varphi(y)$ (henceforth denoted by $g_{m+1}(y)$) and the left-hand sides $g_{j}(y), 1 \leq j \leq m$, of the constraints satisfy the Lipschitz condition
\begin{equation}\label{lipschitz_condition}
	\left|g_j(y')-g_j (y'')\right| \leq L_j \|y'-y''\|, \; y', y''\in D, \; 1\leq j \leq m+1,
\end{equation}
with a priori unknown constants $L_j, 1\leq j \leq m+1$, and may be multiextremal.

@@

	The existence of the $\varepsilon$-reserved solutions of problem (\ref{problem}) can be interpreted as some analog of the regularity conditions in classical nonlinear programming problems. The applied role of this condition should be also noted. Even if the exact solution $y^\ast$ is found, its practical implementation is possible as some approximation of $y^\ast$ only. Therefore, it is important that there are feasible points from $Q_{m+1}$ close enough to $y^\ast$ (in coordinates and values). The existence of the $\varepsilon$-reserved solution guarantees the existence of such points. In this case, the domain $Y_{\epsilon}$ may play the role of a set of suitable approximations.

@@

	Thus, a search trial at some point $x^k \in [0,1]$ performed at the $k^{th}$ iteration of the algorithm will consist of the following sequence of operations:
\begin{itemize} 
  \item determine the image $y^k=y(x^k )$ with the use of the evolvent $y(x)$. 
  \item compute the values $g_1 (y^k ), \ldots ,g_{\nu} (y^k )$, where $\nu = \nu(x^k)$ is the index of the point $x^k$.
	\item fix the result of the trial as three values 
\begin{equation}\label{trial_result}
	x^k, \; z^k = g_{\nu}\left( y(x^k) \right), \; \nu = \nu(x^k).
\end{equation}
\end{itemize}

@@

	In the case when the point $y^k = y(x^k)$ is a feasible one, i.e. when $x^k \in Q_{m+1}$, the trial involves checking all constraints and computing the objective function. Here, the value of the index $\nu$ is accepted to be equal to $m+1$.

	The main idea of the index scheme of accounting for the constraints consists in the reduction of the constrained problem (\ref{one_dimensional_problem}) (the functions of which satisfy the H\"{o}lder condition with corresponding coefficients $K_j$) to an unconstrained problem of the form
\begin{equation}\label{reduction_problem}
	\psi(x^*)=\min \left\{\psi(x): x \in [0,1] \right\},
\end{equation}
\begin{eqnarray} \nonumber
	\psi(x)=\frac{g_{\nu}(y(x))}{K_{\nu}} - 
	\left\{
   \begin{array}{lr}
     0, & \nu \leq m,\\
     \frac{\varphi^\ast}{K_{m+1}}, & \nu = m + 1,
   \end{array}
	\right.
\end{eqnarray}
where $\nu = \nu(x)$ and $\varphi^\ast$ is the objective function value at the point of the problem solution (\ref{one_dimensional_problem}).

@@

\section{Index-Based Global Search Algorithm}
	The index-based global search algorithm under consideration for solving the problem (\ref{problem}) (here it is given according to \cite{Strongin2000}) involves the construction of a sequence of points $x^k \in (0,1)$, at which the search trials are conducted with the results of the type (\ref{trial_result}). The set of triplets 
$\{x^k,\; z^k=g_{\nu}\big(y(x^k)\big),\; \nu=\nu(x^k)\},\; 1 \leq k \leq n,$
makes up the search information accumulated by the method after performing $n$ steps. The operation of the index-based global search algorithm is defined by the following rules.

@@

	The convergence conditions for the index global search algorithm can be formulated in the form of the following theorem from \cite{Strongin2000}. 
\\
\textbf{Theorem 1}. Assume that the following conditions are satisfied:
\begin{enumerate} 
  \item Problem (\ref{problem}) has an $\varepsilon$-reserved solution $y_{\varepsilon}$ from (\ref{epsilon_reserved_solution}).
  \item Functions $g_j(y), 1 \leq j \leq m+1,$ admit Lipschitzian (with the constants $L_j$) extensions $G_{j}(y)$ over the whole feasible domain $D$ from (\ref{D}), i.e.
$$
	g_j \left( y(x) \right) = G_j \left( y(x) \right), \; x \in Q_j, \; 1 \leq j \leq m+1,
$$
where the sets $Q_j$ are from (\ref{Q_intervals}), and $y(x)$ is the curve from (\ref{one_dimensional_problem}).
	\item The parameters of the method $\epsilon_{\nu}, 1 \leq \nu \leq m,$ used in the rule (\ref{z_estimates}) are the corresponding components of the reserve vector $\epsilon_{R}$ from (\ref{epsilon_reserved_solution}).
	\item For a sufficiently large iteration number $k$, the values $\mu_{\nu}$ from (\ref{current_lower_bounds}) satisfy the inequalities
		\begin{equation}\label{theorem_inequalities}
			r_{\nu}\mu_{\nu} > 2^{3-\frac{1}{N}}L_{\nu}\sqrt{N+3}, \; 1 \leq \nu \leq m+1. 
		\end{equation}
Then, any limit point $\bar{y}$ of the sequence of the trials $\{y^k\}=\big\{y\big(x^k\big)\big\}$ generated by the index algorithm for problem (\ref{problem}) belongs to the set $Y_{\epsilon}$ form (\ref{Y_epsilon}) and satisfies the conditions	
$$
	\varphi(\bar y) = \inf\left\{\varphi(y^k):g_j(y^k) \leq 0, 1 \leq j \leq m, k = 1,2,\ldots \right\} \leq \varphi(y_{\epsilon}).
$$
\end{enumerate}
\textbf{Note 1}. The computational scheme of the algorithm is described in the form convenient for  formulating its theoretical properties. The software implementation of this scheme may be made more frugal. For example, according to Step 7, the point $x^{k+1}$ of the next trial will be an internal point of the interval $(x_{t-1}, x_t)$, i.e. $x^{k+1} \in (x_{t-1}, x_t)$. Therefore, the set of trial points $x^1,\ldots, x^k$ (and the related trial results (\ref{trial_result})) may not be ordered  at every step of the algorithm but may be maintained in an ordered state using appropriate data structures.
\\
\textbf{Note 2}. The use of the set of parameters $\epsilon_1,\ldots,\epsilon_m$ (the values of which correspond to the coordinates of the reserve vector $\epsilon_R$) in the index algorithm will significantly reduce the density of the trial sequence in the vicinity of the feasible domain boundaries. However, this raises a question about the particular values of the reserve vector components $\epsilon_R$. In \cite{Strongin2000}, it was proposed to choose the components of the vector $\epsilon_R$ according to the formula 
\begin{equation}\label{epsilon_nu}
	\epsilon_{\nu} = \mu_{\nu}\delta, \; 1 \leq \nu \leq m, 
\end{equation}
where $\mu_{\nu}$ are the current estimates from (\ref{current_lower_bounds}) and $\delta>0$ is a parameter. This approach to choosing the reserve vector components allows one to take into account the properties of different problem functions characterized by their Lipschitz constants. This method for choosing the reserve vector was used in the numerical experiments, the results of which are presented below.
\\
\textbf{Note 3}. The index algorithm for solving the constrained problems is based on the information-statistical algorithm of global search (AGS) for solving unconstrained problems. AGS has been compared repeatedly with both deterministic and nature-inspired global optimization algorithms. The results of the experimental comparison (see, for example, \cite{Sovrasov2019}) show that AGS is superior to many known methods of similar purpose, including DIRECT and DIRECT\textit{l}.

	At the same time, the index scheme of accounting for the constraints also demonstrated its advantages (in the number of computed values of the problem functions) over the traditional penalty functions method in solving a wide range of test problems (see the results of the experimental comparison in \cite{Barkalov2017_1, Barkalov2017_2}.)

@@

	According to Theorem 1, the sequence of trial points $\{y^k\}$ will converge to a point $\bar y \in Y_\epsilon$ if the condition (\ref{theorem_inequalities}) is satisfied. So far, an appropriate choice of the parameters $r_{\nu}$ allows using the values
\begin{equation}\label{estimates_Lipschitz_constants}
	\frac{r_{\nu}\mu_{\nu}}{2^{3-\frac{1}{N}}\sqrt{N+3}}, \; 1 \leq \nu \leq m+1, 
\end{equation}
as the estimates of the Lipschitz constants for the problem functions $g_{j}(y), 1 \leq j \leq m+1$.

	Compliance with condition (\ref{theorem_inequalities}) will be achieved when choosing a high values of the parameters $r_{\nu}$. However, in this case the constant will be overestimated, and the method will perform a large number of trials until the stopping condition is fulfilled.
	The choice of low values of the parameters $r_{\nu}$ (which corresponds to an underestimate of the Lipschitz constant) would reduce the number of trials considerably but could destroy the convergence to the global extremum (due to possible violation of the condition (\ref{theorem_inequalities})).

@@

	Step 5. For each interval $(x_{i-1}, x_i), 1 \leq i \leq k+1$: 
\begin{flushleft}
	\begin{itemize}[\labelitemii]
		\item compute the global characteristics $R_{glob}(i)$ according to formulae (\ref{R_1}), (\ref{R_2}), (\ref{R_3}) using the value $r_{\nu}=r_{\nu}^{glob}$;
		\item compute the local characteristics $R_{loc}(i)$ according to formulae (\ref{R_1}), (\ref{R_2}), (\ref{R_3}) using the value $r_{\nu}=r_{\nu}^{loc}$;
		\item determine the characteristic $R(i)$ as
\begin{equation}\label{R_max}
	R(i) = \max{\left\{ R_{glob}(i), \rho R_{loc}(i) \right\}},
\end{equation}
where $\rho=\Big(\frac{1-1/r_{\nu}^{glob}}{1-1/r_{\nu}^{loc}}\Big)^2$ if $\nu(x_{i-1})=\nu(x_i)=\nu$, and $\rho=1$ if $\nu(x_{i-1}) \neq \nu(x_i)$.
	\end{itemize}
\end{flushleft}

@@

	Let us illustrate the work of the algorithms IA and IA-DL when minimizing the function
\begin{eqnarray}
	\varphi(y_1, y_2)=-1.5y_1^2\exp{\left\{1-y_1^2-20.25(y_1-y_2)^2\right\}}- \nonumber \\
	-\left(0.5(y_1-1)(y_2-1)\right)^4\exp{\left\{2-\left(0.5(y_1-1)\right)^4-(y_2-1)^4\right\}}
	\nonumber
\end{eqnarray}
in the range $0 \leq y_1 \leq 4$, $1 \leq y_2 \leq 3$ subject to the constraints
\begin{eqnarray}
	g_1(y_1, y_2) &=& 0.01 \left( (y_1-2.2)^2+(y_2-1.2)^2-2.25 \right) \leq 0, \nonumber \\
	g_2(y_1, y_2) &=& 100 \left(1-(y_1-2)^2/1.44+(0.5y_2)^2 \right) \leq 0, \nonumber \\
	g_3(y_1, y_2) &=& 10 \left( y_2 - 1.5 - 1.5 \sin{\left( 6.283(y_1-1.75) \right)}\right) \leq 0. \nonumber
\end{eqnarray}
The feasible domain of this problem is non-connected and consists of three subdomains. The global minimum $\varphi(y^\ast)=-1.489$ is achieved at the point $y^\ast =(0.942, 0.944)$ (the values are given with the accuracy of $10^{-3}$).

@@

	In Fig. \ref{fig:example_strongin}, the search domain (square) is shown, the feasible domain is highlighted by green, and the trial points generated by IA (a) and IA-DL (b) are marked. The best point is marked by red. For both methods, the evolvent with the parameter $m=10$ was used, the accuracy of the search for the solution $\epsilon=0.002$; the adaptive reserve parameter $\delta=0.008$ from (\ref{epsilon_nu}). For the IA method, the parameters $r_{\nu}=2.3, 1 \leq \nu \leq 4$, were used. The same values were used for the IA-DL method as the parameters $r_{\nu}^{glob}, 1 \leq \nu \leq 4$. Also, the parameters $r_{\nu}^{loc}=1.5, 1 \leq \nu \leq 4,$ were used. To solve the problem with a predefined accuracy $\epsilon=0.002$, 478 trials were required for the IA method compared to only 303 trials for the IA-DL method. The results of larger-scale numerical experiments are presented in Section 4.

	Let us formulate and prove the convergence conditions for the new algorithm IA-DL.\\
\textbf{Theorem 2}. Let us assume the index algorithm (IA) is applied to solve the problem (\ref{problem}) and for the values $r_{\nu}=r_{\nu}^{glob}, 1 \leq \nu \leq m+1,$ the conditions of Theorem 1 are fulfilled. Then, the algorithm with dual estimate of the Lipschitz constants (IA-DL) with the parameters $r_{\nu}^{glob}>r_{\nu}^{loc}>1,\; 1\leq \nu \leq m+1,$ when solving the same problem also generates a sequence of the trial points $\{y^k\}=\{y(x^k)\}$, any limit point $\bar{y}$ of which belongs to the set $Y_{\epsilon}$ from (\ref{Q_intervals}), i.e. satisfies the condition 
\begin{equation}\label{phi_bar_y}
	\varphi(\bar y) = \inf\left\{\varphi(y^k):g_j(y^k) \leq 0, 1 \leq j \leq m, k = 1,2,\ldots \right\} \leq \varphi(y_{\epsilon}).
\end{equation}

@@

	Since the point $\bar y = y(\bar x)$ belongs to the set $Y_{\epsilon}$, then
$$
	g_{\nu}\left( y(\bar x) \right) \leq -\epsilon_{\nu}, \; 1\leq\nu\leq m,
$$
$$
	g_{m+1}\left( y(\bar x) \right) \leq z_{m+1}^\ast.
$$

	Then, the characteristic $R_{glob}(j)$ of the interval $[x_{j-1}, x_j], j=j(k)$ will be strictly positive, i.e. $R_{glob}(j)>0$. Indeed, at $\nu=\nu(x_j)=\nu(x_{j-1})$ and the value $r_{\nu}=r_{\nu}^{glob}$
\begin{eqnarray}
	R(j) &=& \Delta_i + \frac{(z_j-z_{j-1})^2}{r_{\nu}^2\mu_{\nu}^2\Delta_i} - 2\frac{z_j+z_{j-1}-2z_{\nu}^*}{r_{\nu}\mu_{\nu}}  \nonumber \\
	&\geq& \Delta_i + 4\frac{z_{\nu}^*-\left( g_{\nu}\left( y(\bar x) \right)+L_{\nu}\sqrt{N+3}\left( (x_j-\bar x)^{\frac{1}{N}}+(\bar x - x_{j-1})^{\frac{1}{N}} \right)\right)}{r_{\nu}\mu_{\nu}} \nonumber \\
	&=& \Delta_i-4\frac{L_{\nu}\sqrt{N+3}\left( (x_j-\bar x)^{\frac{1}{N}}+(\bar x - x_{j-1})^{\frac{1}{N}} \right)}{r_{\nu}\mu_{\nu}}+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \nonumber \\
%	&=& \Delta_i-4\frac{L_{\nu}\sqrt{N+3}\left( \alpha^{\frac{1}{N}}+(1-\alpha)^{\frac{1}{N}} \right)}{r_{\nu}\mu_{\nu}}+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \nonumber \\
	&\geq& \Delta_i\left(1-4\frac{L_{\nu}\sqrt{N+3}\max_{0\leq\alpha\leq1} {\left( \alpha^{\frac{1}{N}}+(1-\alpha)^{\frac{1}{N}} \right)}}{r_{\nu}\mu_{\nu}} \right)+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \nonumber \\
	&=& \Delta_i\left(1-4\frac{2^{3-1/N}L_{\nu}\sqrt{N+3}}{r_{\nu}\mu_{\nu}} \right)+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} > 0.
	\nonumber
\end{eqnarray}
At $\nu=\nu(x_j ) > \nu(x_{j-1})$ and the value $r_{\nu}=r_{\nu}^{glob}$
\begin{eqnarray}
	R(j) &=& 2\Delta_i - 4\frac{z_j-z_{\nu}^*}{r_{\nu}\mu_{\nu}} \geq 2\Delta_i+4\frac{z_{\nu}^*-\left( g_{\nu}\left(y(\bar x)\right)+2L_{\nu}\sqrt{N+3}(x_j-\bar x)^{\frac{1}{N}} \right)}{r_{\nu}\mu_{\nu}} \nonumber \\
	&=& 2\Delta_i - 8\frac{L_{\nu}\sqrt{N+3}(x_j-\bar x)^{\frac{1}{N}}}{r_{\nu}\mu_{\nu}}+4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \nonumber \\
	&\geq& 2\Delta_i\left( 1-4\frac{L_{\nu}\sqrt{N+3}}{r_{\nu}\mu_{\nu}} \right) + 4\frac{z_{\nu}^*-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} > 0.
	\nonumber
\end{eqnarray}

@@

\begin{table}[h!]
	\centering
		\begin{tabular}{ccccc}
			\hline
			& $N=2$ & $N=3$ & $N=4$ & $N=5$ \\
			\hline
			IA-DL, $r_{\nu}^{loc}=1.4$ & 316 & 7403 & 40419 & 296357 \\
			\hline
			IA-DL, $r_{\nu}^{loc}=2.0$ & 350 & 8990(1) & 41230 & 360497 \\
			\hline
			IA & 439 & 11804(1) & 41973(3) & 404660(2) \\
			\hline
		\end{tabular}
	\caption{Average number of iterations performed by the methods on “Simple” class}
	\label{tab:average_number_of_iterations_simple}
\end{table}
